---
title: Screening for LSID errors
author: Pieter Provoost
editor_options: 
  chunk_output_type: console
output: rmdformats::robobook
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "docs") })    
---

The goal of this notebook is to screen the OBIS database for LSID issues caused by the auto increment behaviour of Excel and possibly other spreadsheet software.

The first approach we will use is to look for many different LSIDs used for the same `scientificName` within a dataset.

First let's get all `scientificName` and `scientificNameID` combinations by dataset:

```{r error = FALSE, warning = FALSE, message = FALSE}
library(DBI)
library(dplyr)
library(ggplot2)
library(knitr)

if (!file.exists("occurrence.dat")) {
  query <- "
    select
      dataset_id,
      flat->>'scientificName' as scientificName,
      flat->>'scientificNameID' as scientificNameID,
      count(*) as records
    from occurrence
    where flat->>'scientificNameID' is not null
    group by
      dataset_id,
      flat->>'scientificName',
      flat->>'scientificNameID'
  "
  readRenviron("env.txt")
  con <- dbConnect(RPostgres::Postgres(), dbname = Sys.getenv("OBIS_DB"), host = Sys.getenv("OBIS_HOST"), port = Sys.getenv("OBIS_PORT"), user = Sys.getenv("OBIS_USER"), password = Sys.getenv("OBIS_PASSWORD"), bigint = "integer")
  res <- dbSendQuery(con, query)
  occ <- dbFetch(res)
  dbClearResult(res)
  dbDisconnect(con)
  save(occ, file = "occurrence.dat")
} else {
  load("occurrence.dat")
}

head(occ)
```

Now we can calculate by dataset how many `scientificNameID` values there per `scientificName`:

```{r error = FALSE, warning = FALSE, message = FALSE}
stats <- occ %>%
  filter(!is.na(scientificname)) %>%
  group_by(dataset_id, scientificname) %>%
  summarize(ids = length(unique(scientificnameid)), records = sum(records))

stats
```

Let's visualize:

```{r error = FALSE, warning = FALSE, message = FALSE}
ggplot(stats) +
  geom_histogram(aes(ids), binwidth = 1) +
  scale_y_continuous(trans = "log10")
```

Now check the datasets with 5 or more identifiers per name:

```{r error = FALSE, warning = FALSE, message = FALSE}
stats_nok <- stats %>%
  filter(ids >= 5)

stats_nok %>%
  rmarkdown::paged_table()
```

Then get the problematic datasets:

```{r error = FALSE, warning = FALSE, message = FALSE}
datasets <- robis::dataset() %>%
  select(id, url)

stats_nok %>%
  group_by(dataset_id) %>%
  summarize(species_nok = n(), records = sum(records)) %>%
  left_join(datasets, by = c("dataset_id" = "id")) %>%
  arrange(desc(records)) %>%
  kable()
```